analysis

### 2024-12-08 analysis.qmd
added q2 code

```{r, echo= FALSE}
library("dplyr")
library(mice)
library(sjlabelled)
library(tidyverse)

@@

```{r}
#Need to convert school funding adequacy imputed values to school funding categorical variable before using in models
# imputated data has 5 dataframes within a list, need to run a loop for all five datasets to create a categorical variable for each set. 
```

@@

```{r}
#Variance Inflaction Factor (VIF)
library(car)
vif(mlrmodel)

@@

```{r}
#Logistic 
lmod <- glm(School.Funding.Cat ~ Residential.Segregation.Index + Median.Household.Income + Median.Prop.Tax,
              data=combined,
              family="binomial")
summary(lmod)

with(sub.imp, glm(School.Funding.Cat ~ Residential.Segregation.Index + Median.Household.Income + Median.Prop.Tax, family="binomial"))

#Outcome = School.Funding.Cat
#Predictors = residential segregation, median.household.income, median.prop.tax
```

>>>>

```{r, echo= FALSE, warning = FALSE}
library("dplyr")
library(mice)
library(sjlabelled)
library(tidyverse)

@@

```{r}
#Need to convert school funding adequacy imputed values to school funding categorical variable before using in models
# Imputated data has 5 dataframes within a list, need to run a loop for all five datasets to create a categorical variable for each set.

sub <- combined |> select(c("Life.Expectancy","Residential.Segregation.Index","X..Limited.Access.to.Healthy.Foods","X..Uninsured.Adults","Median.Household.Income", "School.Funding.Adequacy", "Median.Prop.Tax", "Region_fac"))
sub <- unlabel(sub) #unlabel the data (labels cause problem for the mice function)

# Perform the imputation
sub$School.Funding.Cat <- ifelse(sub$School.Funding.Adequacy > 0, 1, 0)  # Add the new variable to the dataset
sub.imp <- mice(sub, m = 5, method = "pmm", print = FALSE)

```{r}
# To create a School.Funding.Categorical variable, create a function and run the loop

# Define the categorization function
categorize_school_funding <- function(x) {
  ifelse(x > 0, 1, 0)  # Binary classification
}

# Loop over the imputations
for (i in 1:5) {
  # Extract the imputed values for the i-th dataset
  imputed_values <- sub.imp$imp$School.Funding.Adequacy[, i]

  # Categorize the imputed values
  categorized_values <- categorize_school_funding(imputed_values)
  complete_cat <- rep(NA, nrow(sub))  # Initialize with NAs

  # Get the indices where imputation occurred
  imp_indices <- which(!is.na(categorized_values))  # Only for imputed values

  # Assign the categorized values to the rows that were imputed
  complete_cat[imp_indices] <- categorized_values[imp_indices]

  # Save into the `imp` list of the `sub.imp` object
  sub.imp$imp$School.Funding.Cat[[i]] <- complete_cat[imp_indices]
}
```

```{r}
long_data <- complete(sub.imp, action = "long", include = TRUE)  # Include original data

# Check the structure
str(long_data)

# Ensure that `School.Funding.Cat` is correctly populated
long_data$School.Funding.Cat <- ifelse(long_data$School.Funding.Adequacy > 0, 1, 0)

if (!".imp" %in% colnames(long_data) || !".id" %in% colnames(long_data)) {
  stop("`.imp` or `.id` columns are missing in the long data!")
}
```

@@

```{r}
#Variance Inflaction Factor (VIF)
library(car)
vif(mlrmodel)
```

@@

```{r}
## Checking to see if the imputated data has a large effect on the outcome

# Logistic Regression without Imputated Data
# Outcome = School.Funding.Cat
# Predictors = residential segregation, median.household.income, median.prop.tax

lmod <- glm(School.Funding.Cat ~ Residential.Segregation.Index + Median.Household.Income + Median.Prop.Tax,
              data=combined,
              family="binomial")

library("modelsummary")
modelsummary(lmod,
             fmt = fmt_significant(2),
             shape = term ~ model + statistic,
             statistic = c("std.error","conf.int","p.value"),
             exponentiate = FALSE,
             gof_map=NA)

modelsummary(lmod,
             fmt = fmt_significant(2),
             shape = term ~ model + statistic,
             statistic = c("std.error","conf.int","p.value"),
             exponentiate = TRUE,
             gof_map=NA)

```{r}
imp_mods <- with(sub_imp, glm(School.Funding.Cat ~ Residential.Segregation.Index + Median.Household.Income + Median.Prop.Tax), family="binomial")

modelsummary(pool(imp_mods),
             fmt = fmt_significant(2),
             shape = term ~ model + statistic,
             statistic = c("std.error","conf.int","p.value"),
             exponentiate = FALSE,
             gof_map=NA)

modelsummary(pool(imp_mods),
             fmt = fmt_significant(2),
             shape = term ~ model + statistic,
             statistic = c("std.error","conf.int","p.value"),
             exponentiate = TRUE,
             gof_map=NA)
```

```{r}
# Creating the AUC curve

library(caret)
library(pROC)

set.seed(42)
train_idx <- sample(1:nrow(long_data), size = 0.7 * nrow(long_data))
train_data <- long_data[train_idx, ]
test_data <- long_data[-train_idx, ]

glm_model <- glm(School.Funding.Cat ~ Residential.Segregation.Index + Median.Household.Income + Median.Prop.Tax, data = train_data, family = binomial)

# Step 4: Make predictions on the test set
test_probs <- predict(glm_model, newdata = test_data, type = "response")
test_pred <- ifelse(test_probs > 0.5, 1, 0)  # Classify based on threshold 0.5

# Step 5: Create a confusion matrix
conf_matrix <- confusionMatrix(factor(test_pred), factor(test_data$School.Funding.Cat))
print(conf_matrix)
```
